### 2018-04-11 paper/background.tex

\subsubsection{Structure}
% definition of CHC
As stated in \autoref{sec:intro} a Constrained Horn Clause is a
logical implication where the antecedent is called the body and the
consequent is called the head.
%
The body is a conjunction of a logical formula, called the fact,
and a vector of uninterpreted predicates. The fact is an arbitrary
formula in some background logic, such as linear integer arithmetic.
%
The uninterpreted predicates are applied to variables which may or may
not appear in the fact.
%
A head can be either an uninterpreted predicate applied to variables or \cc{False}.
%
A Clause where the head is \cc{False} is called query.
\begin{defn}
  CHCs can be defined structurally:
\begin{align*}
  \cc{chc} &\Coloneqq \cc{head} \gets \cc{body} \\
  \cc{head} &\Coloneqq \cc{False} \\&\mid \cc{pred}  \\
  \cc{body} &\Coloneqq \varphi \wedge \cc{preds} \\
  \cc{preds} &\Coloneqq \cc{True} \\&\mid \cc{pred} \wedge \cc{preds} \\
  \cc{pred} &\Coloneqq \textsf{an uninterpreted predicate applied to variables} \\
  \varphi &\Coloneqq \textsf{a formula that all variables are vocabulary} \\
\end{align*}
\end{defn}
%
% define a body function, head function, and recursion-free
For a given CHC $C$, $\bodyof{C}$ denotes the vector of uninterpreted
predicates in the body and $\consof{C}$ denotes the fact in the
body.
%
It is possible for an uninterpreted predicate to appear multiple times
in the body of $C$.
%
If $C$ is not a query, then $\headof{C}$ denotes the uninterpreted
predicate in the head.
%

@@

%define siblings
\begin{defn}
  Given a CHC system $\cc{S}$ and two uninterpreted predicates
  $\cc{P}$ and $\cc{Q} \in \predof{S}$, if $\exists \cc{C} \in \cc{S}$
  such that $\cc{P} \in \bodyof{C}$ and $\cc{Q} \in \bodyof{C}$, then
  $\cc{Q}$ and $\cc{P}$ are siblings.
\end{defn}
%
\begin{ex}
  Because uninterpreted predicates $\cc{L}_9$ and $\cc{db}$ both
  appear in the body of clause (6), $\cc{L}_9$ and $\cc{db}$ are
  siblings.
  %
\end{ex}
%
Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
$\siblingof{P}$ denotes the set of all predicate siblings of $\cc{P}$
in $\cc{S}$.
%
For a given CHC system $\cc{S}$, if there are no uninterpreted
predicate $\cc{P} \in \predof{S}$ such that $\cc{P} \in \tdepsof{P}$,
then $\cc{S}$ is a \emph{recursion-free} CHCs system.
%

@@

% interpolants:
\subsection{Logical interpolation}
\label{sec:itps}
%
\sys solves CHC systems, logic-programming problems in which a
solution maps predicate symbols to first-order formulas.
%
All objects defined in this paper are defined over a fixed space of
first-order logical variables $X$.
%
In particular, for theory $\mathcal{T}$, the space of $\mathcal{T}$
formulas over $X$ is denoted $\tformulas{\mathcal{T}}$.
%
For each formula $\varphi \in \tformulas{\mathcal{T}}$, the set of
variables that occur in $\varphi$ (i.e., the \emph{vocabulary} of
$\varphi$) is denoted $\vocab(\varphi)$.
% define models, satisfaction, entailment
For formulas $\varphi_0, \ldots, \varphi_n, \varphi \in
\tformulas{\mathcal{T}}$, the fact that $\varphi_0, \ldots, \varphi_n$
\emph{entail} $\varphi$ is denoted $\varphi_0, \ldots, \varphi_n
\entails \varphi$.

@@

% introduce interpolation:
An interpolant of mutually inconsistent $\mathcal{T}$ formulas
$\varphi_0$ and $\varphi_1$ is a $\mathcal{T}$ formula in their common
vocabulary that explains their inconsistency.
%
\begin{defn}
  \label{defn:itps}
  % define:
  For $\varphi_0, \varphi_1 \in \tformulas{\mathcal{T}}$, $I \in
  \tformulas{\mathcal{T}}$ such that
  %
  \textbf{(1)} $\varphi_0 \entails I$, %
  \textbf{(2)} $I, \varphi_1 \entails \false$, and %
  \textbf{(3)} $\vocab(I) \subseteq \vocab(\varphi_0) \intersection
  \vocab(\varphi_1)$,
  %
  $I$ is an \emph{interpolant} of $\varphi_0$ and $\varphi_1$.
\end{defn}
%
All spaces of formulas in the remainder of this paper will be defined
for a fixed, arbitrary theory $\mathcal{T}$ that supports
interpolation, such as combinations of the theories of linear
arithmetic and the theory of uninterpreted functions with equality.
% introduce decision procedure:
Although determining the satisfiability of formulas in such theories
is NP-complete in general, decision procedures~\cite{moura08} and
interpolating theorem provers~\cite{mcmillan04} for such theories have
been proposed that operate on such formulas efficiently.
%
\sys is defined as using a decision procedure for $\mathcal{T}$ named
\issat, and an interpolating theorem prover for $\mathcal{T}$ named
$\solveitp$.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "p"
%%% End:

>>>>

\subsubsection{Structure}
% definition of CHC
A Constrained Horn Clause is a logical implication where the
antecedent is called the body and the consequent is called the head.
%
The body is a conjunction of a logical formula, called the fact,
and a vector of uninterpreted predicates. The fact is an arbitrary
formula in some background logic, such as linear integer arithmetic.
%
The uninterpreted predicates are applied to variables which may or may
not appear in the fact.
%
A head can be either an uninterpreted predicate applied to variables or \cc{False}.
%
A clause where the head is \cc{False} is called a query. A CHC can be
defined structurally:
\begin{align*}
\cc{chc} &\Coloneqq \cc{head} \gets \cc{body} \\
\cc{head} &\Coloneqq \cc{False} \\&\mid \cc{pred}  \\
\cc{body} &\Coloneqq \varphi \wedge \cc{preds} \\
\cc{preds} &\Coloneqq \cc{True} \\&\mid \cc{pred} \wedge \cc{preds} \\
\cc{pred} &\Coloneqq \textsf{an uninterpreted predicate applied to variables} \\
\varphi &\Coloneqq \textsf{a formula that all variables are vocabulary} \\
\end{align*}
%
% define a body function, head function, and recursion-free
For a given CHC $C$, $\bodyof{C}$ denotes the vector of uninterpreted
predicates in the body and $\consof{C}$ denotes the fact in the
body.
%
It is possible for an uninterpreted predicate to appear multiple times
in the body of $C$.
%
If $C$ is not a query, then $\headof{C}$ denotes the uninterpreted
predicate in the head.
%

@@

%define siblings
\begin{defn}
  Given a CHC system $\cc{S}$ and two uninterpreted predicates
  $\cc{P}$ and $\cc{Q} \in \predof{S}$, if $\exists \cc{C} \in \cc{S}$
  such that $\cc{P} \in \bodyof{C}$ and $\cc{Q} \in \bodyof{C}$, then
  $\cc{Q}$ and $\cc{P}$ are siblings.
\end{defn}
%
\begin{ex}
  Because uninterpreted predicates $\cc{L}_9$ and $\cc{db}$ both
  appear in the body of clause (6), $\cc{L}_9$ and $\cc{db}$ are
  siblings.
  %
\end{ex}
%
Given a CHC system $\cc{S}$ and an uninterpreted predicate $\cc{P}$,
$\siblingof{P}$ denotes the set of all siblings of $\cc{P}$
in $\cc{S}$.
%
For a given CHC system $\cc{S}$, if there are no uninterpreted
predicate $\cc{P} \in \predof{S}$ such that $\cc{P} \in \tdepsof{P}$,
then $\cc{S}$ is a \emph{recursion-free} CHCs system.
%

@@

% interpolants:
\subsection{Logical interpolation}
\label{sec:itps}
%
All objects defined in this paper are defined over a fixed space of
first-order logical variables $X$.
%
In particular, for a theory $\mathcal{T}$, the space of $\mathcal{T}$
formulas over $X$ is denoted $\tformulas{\mathcal{T}}$.
%
For each formula $\varphi \in \tformulas{\mathcal{T}}$, the set of
variables that occur in $\varphi$ (i.e., the \emph{vocabulary} of
$\varphi$) is denoted $\vocab(\varphi)$.
%
% define models, satisfaction, entailment
For formulas $\varphi_0, \ldots, \varphi_n, \varphi \in
\tformulas{\mathcal{T}}$, the fact that $\varphi_0, \ldots, \varphi_n$
\emph{entail} $\varphi$ is denoted $\varphi_0, \ldots, \varphi_n
\entails \varphi$.

@@

% introduce interpolation:
An interpolant of mutually inconsistent $\mathcal{T}$ formulas
$\varphi_0$ and $\varphi_1$ is a $\mathcal{T}$ formula in their common
vocabulary that explains their inconsistency.
%
\begin{defn}
  \label{defn:itps}
  % define:
  For $\varphi_0, \varphi_1, I \in \tformulas{\mathcal{T}}$, if
  %
  \textbf{(1)} $\varphi_0 \entails I$, %
  \textbf{(2)} $I \land \varphi_1 \entails \false$, and %
  \textbf{(3)} $\vocab(I) \subseteq \vocab(\varphi_0) \intersection
  \vocab(\varphi_1)$,
  %
  then $I$ is an \emph{interpolant} of $\varphi_0$ and $\varphi_1$.
\end{defn}
%
For the remainder of this paper, all spaces of formulas will be
defined for a fixed, arbitrary theory $\mathcal{T}$ that supports
interpolation, such as combinations of the theories of linear
arithmetic and the theory of uninterpreted functions with equality.
% introduce decision procedure:
Although determining the satisfiability of formulas in such theories
is NP-complete in general, decision procedures~\cite{moura08} and
interpolating theorem provers~\cite{mcmillan04} for such theories have
been proposed that operate on such formulas efficiently.
%
We define \sys in terms of an abstract decision procedure for
$\mathcal{T}$ named \issat, and an abstract interpolating theorem
prover for $\mathcal{T}$ named $\solveitp$.

### 2018-04-11 paper/introduction.tex

% introduce the problem of solving recursion-free systems:
In this work we focus on the subclass of CHC systems which are known
as \emph{recursion-free}. In a recursion-free CHC system, the
derivation of each predicate will not contain
itself. \QZ{\DAH{I still think we need a small example here}}
%
Recursion-free CHC systems are an important subclass for two reasons.
%
First, recursion-free systems can be used to model safety properties
for hierarchical programs~\cite{lal-qadeer15,lal-qadeer-lahiri12}
(programs with only bounded iteration and recursion).

%
Second and most importantly, A well-known approach for solving a general CHC
system can reduce to solving a sequence of recursion-free
systems \QZ{reference}.
%
Many solvers attempt to synthesize the solution of the original system 
by using the solutions of constructed recruion-free system~\cite{bjorner13}.
%
Because of this, the performance of solving general CHC system is heavily reliant on the 
performance of the solving recursion-free CHC system.
%

% current general techniques for solving recursion-free systems
Typically, recursion-free CHC systems are not solved directly.
%
Instead, they are reduced to a more specific subclass of
recursion-free CHC system.
%
These classes include those of
\emph{body disjoint} (or \emph{derivation tree})
systems~\cite{heizmann10,bjorner13,mcmillan14,rummer13a,rummer13b} and
of \emph{linear} systems~\cite{albarghouthi12a}.
%
We will discuss these classes in \autoref{sec:overview} and
\autoref{sec:related-work}.
%
Such classes can be solved by issuing
\emph{interpolation queries} for uninterpreted predicate and 
each constrains shows only one time in the interpolation query.
%
We
refer such class as \emph{directly solvable}.

% Time complexity
In general, solving a recursion-free CHC system for 
propositional logic and theory of linear integer arithmetic is
co-NEXPTIME-complete~\cite{rummer13b}.
%
In contrast, solving a directly solvable system with the same logic and theories
is in co-NP~\cite{rummer13b}.
%
Therefore, the performance of solving a recurion-free system is highly reliant on the
size of the directly solvable systems it is reduced to.

@@

% a new solver
The second contribution of this paper is a solver for CHC systems,
named \sys.
%
Given a recursion-free system $S$, \sys reduces the problem
of solving $S$ to solving an equivalent CDD system
$S'$.
%
In the worst case, it is possible that the size of $S'$ may be
exponential in the size of $S$.
%
However, empirically we have found that the size of $S'$ is usually
close enough to the size of $S$ that \sys often outperforms the best
known CHC solvers.
%
The procedure implemented in \sys is a generalization of existing
techniques that synthesize compact verification conditions for
hierarchical programs~\cite{flanagan01,lal-qadeer15}.
%
Given a general (possibly recursive) CHC system, \sys solves a
sequence of recursion-free systems.
%
Each subsystem is a bounded unwinding of the original system. \sys
attempts to combine the solutions of these recursion-free systems to
synthesize a soltuion to the original problem, as has been proposed
in previous work~\cite{rummer13b}.

@@

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "p"
%%% End:

>>>>

% introduce the problem of solving recursion-free systems:
In this work we focus on the subclass of CHC systems which are known
as \emph{recursion-free}. In a recursion-free CHC system, the
derivation of each predicate will not contain
itself. \QZ{\DAH{I still think we need a small example here}}
%
Recursion-free CHC systems are an important subclass for two reasons.
%
First, recursion-free systems can be used to model safety properties
for hierarchical programs~\cite{lal-qadeer15,lal-qadeer-lahiri12}
(programs with only bounded iteration and recursion).
%
Second and most importantly, a well-known approach for solving a
general CHC system can reduces the input problem to solving a sequence
of recursion-free systems \QZ{reference}.
%
Such approaches attempt to synthesize a solution for the   original
system from the solutions of recursion-free systems~\cite{bjorner13}.
%
Because of this, the performance of solving general CHC system relies
heavily on the performance of solving recursion-free CHC systems.
%

% current general techniques for solving recursion-free systems
Typically, recursion-free CHC systems are not solved directly.
%
Instead, they are reduced to a more specific subclass of
recursion-free CHC system.
%
These classes include those of
\emph{body disjoint} (or \emph{derivation tree})
systems~\cite{heizmann10,bjorner13,mcmillan14,rummer13a,rummer13b} and
of \emph{linear} systems~\cite{albarghouthi12a}.
%
We will discuss these classes in \autoref{sec:overview} and
\autoref{sec:related-work}.
%
Such classes can be solved by issuing
\emph{interpolation queries} to find suitable definitions for the
uninterpreted predicates.
%
We refer such classes as \emph{directly solvable}.

% Time complexity
In general, solving a recursion-free CHC system for
propositional logic and the theory of linear integer arithmetic is
co-NEXPTIME-complete~\cite{rummer13b}.
%
In contrast, solving a directly solvable system with the same logic
and theories is in co-NP~\cite{rummer13b}.
%
Therefore, the performance of solving a recursion-free system is
highly reliant on the size of the directly solvable system it reduces
to.

@@

% a new solver
The second contribution of this paper is a solver for CHC systems,
named \sys.
%
Given a recursion-free system $S$, \sys reduces the problem
of solving $S$ to solving an equivalent CDD system
$S'$.
%
In the worst case, it is possible that the size of $S'$ may be
exponential in the size of $S$.
%
However, empirically we have found that the size of $S'$ is usually
close enough to the size of $S$ that \sys often outperforms the best
known CHC solvers.
%
The procedure implemented in \sys is a generalization of existing
techniques that synthesize compact verification conditions for
hierarchical programs~\cite{flanagan01,lal-qadeer15}.
%
Given a general (possibly recursive) CHC system, \sys solves a
sequence of recursion-free systems.
%
Each subsystem is a bounded unwinding of the original system. \sys
attempts to combine the solutions of these recursion-free systems to
synthesize a solution to the original problem, as has been proposed
in previous work~\cite{rummer13b}.

@@
